Certified Knowledge Compilation \\ with Application to Verified Model Counting

### 2023-03-17 doc/mc-paper.tex
create Lean appendix and shorten section
Bug fixes in MICE
et al.

One approach to model counting, known as {\em knowledge compilation},
transforms the formula into a structured form for which model counting
becomes straightforward.  For example, the {\em deterministic,
  decomposable, negation normal form} (d-DNNF) introduced by
Darwiche~\cite{darwiche:aaai:2002,darwiche:ecai:2004} represents a
Boolean formula as a directed acyclic graph, with terminal nodes
labeled by Boolean variables and their complements, and with each
nonterminal node labeled by Boolean operation And or Or.  Restrictions
are placed on the structure of the graph such that a count of the
models can be computed by a single bottom-up traversal.
Kimmig, et al., present a very general framework, known as {\em
  algebraic model counting}~\cite{kimmig:jal:2017}, describing
properties of Boolean functions that can be efficiently computing from
a d-DNNF representation.  These include standard and weighted model
counting, and much more.

@@

We encountered a significant number of errors running their tools.
These seem mostly to be easily fixable, and we are currently
supporting their debugging efforts.  At present, we can summarize the results as:
\begin{itemize}
\item 84 total data points
\item 67 were completed by both programs in under 1000 seconds
\item Of those 21 were faster with MICE, 46 with CRAT
\item 7 completed with CRAT under 1000 seconds but exceeded that threshold for MICE
\item 1 completed with MICE under 1000 seconds but exceeded that threshold for CRAT
\item 9 completed with CRAT but generated errors for MICE.
\end{itemize}

@@

>>>>

One approach to model counting, known as {\em knowledge compilation},
transforms the formula into a structured form for which model counting
becomes straightforward.  For example, the {\em deterministic,
  decomposable, negation normal form} (d-DNNF) introduced by
Darwiche~\cite{darwiche:aaai:2002,darwiche:ecai:2004} represents a
Boolean formula as a directed acyclic graph, with terminal nodes
labeled by Boolean variables and their complements, and with each
nonterminal node labeled by Boolean operation And or Or.  Restrictions
are placed on the structure of the graph such that a count of the
models can be computed by a single bottom-up traversal.
Kimmig et al.\ present a very general framework, known as {\em
  algebraic model counting}~\cite{kimmig:jal:2017}, describing
properties of Boolean functions that can be efficiently computing from
a d-DNNF representation.  These include standard and weighted model
counting, and much more.

@@

We encountered a significant number of errors running their tools.
These seem mostly to be easily fixable, and we are currently
supporting their debugging efforts.  At present, we can summarize the results as:
\begin{itemize}
\item 84 total data points
\item 76 were completed by both programs in under 1000 seconds
\item Of those 21 were faster with MICE, 55 with CRAT
\item 7 completed with CRAT under 1000 seconds but exceeded that threshold for MICE
\item 1 completed with MICE under 1000 seconds but exceeded that threshold for CRAT
\end{itemize}

@@

\input{leanappendix}
